Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-06-13 thirdPaper/draft/typeConeAssociahedra.tex
minor typo changes

The associahedron is a ``mythical polytope''~\cite{Haiman} whose facial structure encodes Catalan families: its vertices correspond to parenthesizations of a non-associative product, triangulations of a convex polygon, or binary trees; its edges correspond to applications of the associativity rule, diagonal flips, or edge rotations; and in general its faces correspond to partial parenthesizations, diagonal dissections, or Schr\"oder trees.
Its combinatorial and topological structure was introduced in early works of D.~Tamari~\cite{Tamari} and J.~Stasheff~\cite{Stasheff}, and a first $3$-dimensional polytopal model was realized by J.~Milnor for the PhD defense of J.~Stasheff.
The first systematic polytopal realizations were constructed by M.~Haiman~\cite{Haiman} and C.~Lee~\cite{Lee}.
Since then, the associahedron has been largely ``demystified'' with several polytopal constructions and generalizations (some are discussed below).
Its realizations can be classified into three species: the secondary polytopes~\cite{GelfandKapranovZelevinsky, BilleraFillimanSturmfels}, the $\b{g}$-vectors realizations~\cite{Loday, HohlwegLange, LangePilaud, HohlwegLangeThomas, HohlwegPilaudStella, Postnikov, PilaudSantos-brickPolytope, PilaudStump-brickPolytope} and the $\b{d}$-vector realizations~\cite{ChapotonFominZelevinsky, CeballosSantosZiegler}. See~\cite{CeballosSantosZiegler} for a discussion of these realizations and their connections.

@@

This paper focuses on a surprising construction of the associahedron that recently appeared in mathematical physics.
Motivated by the prediction of the behavior of scattering particules, N.~Arkani-Hamed, Y.~Bai, S.~He, and G.~Yan recently described in~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} the kinetic associahedron.
It is a class of polytopal realizations of the classical associahedron obtained as sections of a high-dimensional positive orthant with well-chosen affine subspaces.
This construction provides a large degree of freedom in the choice of the parameters defining these affine subspaces, and actually produces all polytopes whose normal fan is affinely equivalent to that of J.-L.~Loday's associahedron~\cite{Loday} (see \cref{subsec:typeConeAsso}).
These realizations were then extended by V.~Bazier-Matte, G.~Douville, K.~Mousavand, H.~Thomas and E.~Y\i ld\i r\i m~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the context of finite type cluster algebras using tools from representation theory of quivers.
More precisely, they fix a finite type cluster algebra~$\cal{A}$ and consider the real euclidean space~$\R^\cal{V}$ indexed by the set~$\cal{V}$ of cluster variables of~$\cal{A}$.
Starting from an acyclic initial seed~$\seed_\circ$, they consider a set~$\cal{M}$ of mesh mutations and compute the corresponding linear dependences between the $\b{g}$-vectors, which can be interpreted as linear spaces in~$\R^\cal{V}$.
Finally, they perturb these mesh linear spaces by a collection~$\b{\ell} \in \R^{\cal{M}}$ of positive parameters and intersect the resulting perturbed mesh affine spaces with the positive orthant of~$\R^\cal{V}$ (see \cref{subsec:typeConeCA}).
They show that the resulting polytope is always a generalized associahedron, and that its normal fan is the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$.
An implicit by-product of their construction is that the space of all polytopal realizations of the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$ is a simplicial cone.

\medskip
In this paper, we revisit, extend and explore further this construction using a reversed approach.
Given a complete simplicial fan~$\Fan$, we consider the space~$\typeCone(\Fan)$ of all its polytopal realizations.
This space was called type cone of~$\Fan$ in~\cite{McMullen-typeCone} and deformation cone of~$\Fan$ in~\cite{Postnikov, PostnikovReinerWilliams}, who studied the case when~$\Fan$ is the braid arrangement leading to the rich theory of generalized permutahedra (recently extended to all finite Coxeter arrangements in~\cite{ACEP-DeformationsCoxeterPermutahedra}).
The type cone is known to be a polyhedral cone defined by a collection of inequalities corresponding to the linear dependences among the rays of~$\Fan$ contained in pairs of adjacent maximal cones of~$\Fan$ (see \cref{def:typeCone}).
Our approach is based on an elementary but powerful observation: for any fan~$\Fan$, all polytopal realizations of~$\Fan$ can be described as sections of a high dimensional positive orthant with a collection of affine subspaces parametrized by the type cone~$\typeCone(\Fan)$ (see \cref{prop:alternativePolytopalRealization}); if moreover the type cone~$\typeCone(\Fan)$ is a simplicial cone, it leads to a simple parametrization of all polytopal realizations of~$\Fan$ by a positive orthant corresponding to the facets of the type cone~$\typeCone(\Fan)$ (see \cref{coro:simplicialTypeCone}).
To prove that the type cone~$\typeCone(\Fan)$ is simplicial, we just need to identify which pairs of adjacent maximal cones of~$\Fan$ correspond to the facets of~$\typeCone(\Fan)$ and to show that the corresponding linear dependences among their rays positively span the linear dependence among the rays of any pair of adjacent maximal cones of~$\Fan$.
When applied to the $\b{g}$-vector fans of cluster algebras (see \cref{subsec:typeConeCA}), this reversed approach yields all polytopal realizations of the $\b{g}$-vector fans and thus efficiently revisits and extends the construction of~\cite{BazierMatteDouvilleMousavandThomasYildirim}.
This new perspective has several advantages, as our proof uniformly~applies~to:

@@

In this section, we apply the results of \cref{subsect::Statement cluster cats} in order to prove \cref{prop:uniqueExchangePropertyCA} and \cref{prop:meshMutations} of \cref{subsec:typeConeCA}.

We first recall the results on categorification of cluster algebras that we will make use of in the proofs.
\begin{theorem}
\label{thm::categorification}
Let~$\cat$ be the cluster category of a valued quiver of Dynkin type~$A,B,C,D,E,F$ or~$G$. Let~$T$ be a basic cluster-tilting object in~$\cat$, and let~$Q$ be the valued Gabriel quiver of~$\End{\cat}(T)$. Then there is a bijection
\[
\CC: \ind(\cat) \xrightarrow{} \{\textrm{cluster variables in the cluster algebra of~$Q$}\},
\]
expressing cluster variables as Laurent polynomials in the variables of the cluster~$\CC(T)$ and which has the follwing properties.
\begin{itemize}
\item[(i)] \emph{(}\cite[Prop.~3.2]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are compatible if and only if~$\cat(X, \susp Y) = 0$.
\item[(ii)] \emph{(}\cite[Prop.~4.3]{FuKeller} and \cite[Prop.~2.2]{Palu}, where the proofs also work over an arbitrary base field\emph{)} For any~$X \in \ind(\cat)$, the~$\b{g}$-vector of~$\CC(X)$ is~$\b{g}([X])$, where we identify~$\Z^n$ and~$K_0(\cat ; T)$ via the isomorphism sending~$(a_1, \dots, a_n)$ to~$\sum_{i=1}^n a_i [T_i]$.
\item[(iii)] \emph{(}\cite[Thm.~7.5]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are exchangeable if and only if~$\dim_{\field_X}\cat(X, \susp Y) = \dim_{\field_Y}\cat(X, \susp Y) = 1$.
\item[(iv)]\emph{(}\cite{BuanIyamaReitenScott} and \cite[Thm.~5.1]{BuanMarshReiten-mutation}, adapted to the cluster categories of species of Dynkin type\emph{)} Let~$T'$ be any basic cluster-tilting object of~$\cat$. Then the matrix~$B'$ of the associated seed~$(\CC(T'),B')$ is given by the multiplicities of the middle terms of the exchange triangles. 
\end{itemize}
\end{theorem}

@@

\begin{proof}
 If~$(x,x')$ is a positive mesh mutation, then~$\{x,x'\}$ is in~$\meshes$.  Indeed, if~$X$ and~$X'$ are the corresponding objects in~$\cat$, then there is an almost-split triangle
 \[
  X\to E\to X' \to \susp E,
 \]
 so that~$X' = \tau^{-1}X$.  Moreover, there is a slice of the Auslander--Reiten quiver of~$\cat$ containing~$X$ and all indecomposable direct factors of~$E$.  The direct sum of indecomposable objects in this slice is a cluster-tilting object in which~$X$ is a source, and mutation at~$X$ gives~$X'$.  Thus~$\{x,x'\}$ is in~$\meshes$.  This gives an injective map from the set of positive mesh mutations to~$\meshes$.

 \smallskip

 Next, assume that~$\{x,x'\}\in \meshes$ starts in~$x$.  Then there is a cluster-tilting object~$T$ in~$\cat$ having~$X$ as a direct factor, and such that~$X$ is a source in~$T$ and mutation at~$X$ yields~$X'$.  Since~$X$ is a source, the mutation triangle starting and ending in~$X$ are
 \[
  X\to E\to X' \to \susp X \quad \textrm{and} \quad  X' \to 0 \to X \to \susp X'.
 \]
  Therefore~$X' = \susp^{-1}X = \tau^{-1}X$.  Thus the first triangle is the almost-split triangle starting in~$X$, since the dimension of~$\cat(X',\susp X)$ is~$1$ over~$\field_{X'}$.  Since the pair~$\{x,x'\}$ is not initial,~$\phi(X')$ is not an initial variable; in other words,~$X'$ is not in~$\add(T)$.  Therefore,~$X = \tau X' = \susp X'$ is not in~$\add\susp T$, so the mutation is a positive mesh mutation by Lemma~\ref{lemm::almost-split}.  This gives an injective map from~$\meshes$ to the set of positive mutations. 
\end{proof}

\begin{corollary}[\cref{lem:bijectionMeshMutations}]
\label{coro::number of positive meshes}
 We have that~$|\meshes|$ is the number of non-initial cluster variables.
\end{corollary}

@@

\begin{proof}
In view of \cref{thm::categorification} (iii), it follows from~\cite[Lem.~3.3]{Palu}.
\end{proof}

@@

\begin{proof}
Let~$\cat$ be a cluster category of Dynkin type~$A,B,C,D,E,F$ or~$G$, and let $T\in\cat$ be some basic cluster-tilting object categorifying the given initial exchange matrix~$\B_\circ$. 
By \cref{lemma:positive mutation triangles}, we may assume that the mutation under consideration is a positive mutation.
In view of \cref{thm::categorification} (ii) and (iv) it is enough to show that the relation in~$K_0(\cat ; T)$ given by any triangle ${X\to E\to Y\xrightarrow{h} \susp X}$ with~$h\in(\susp T)$ is a positive linear combination of relations coming from Auslander--Reiten triangles with third term not in~$\add(\susp T)$.
This categorified statement is precisely \cref{corollary::meshes positively generate cluster cats}.
\end{proof}

It also follows form \cref{thm::categorification} above that any cluster algebra of finite type has the unique exchange relation property (see \cref{def:uerp}):
\begin{corollary}
\label{corollary:UERPforCAproof}
\emph{(\cref{prop:uniqueExchangePropertyCA})}
Let~$\B_\circ$ be any finite type exchange matrix, and let~$\mathcal{A}(\B_\circ)$ be the associated cluster algebra without coefficients.
Then, for any exchangeable cluster variables~$x$ and~$x'$, the linear dependence $\sum_{y \in \cluster \cup \cluster'} \coefficient[{\b{g}_y}][\cluster][\cluster'] \, \b{g}_y = 0$ only depends on the pair~$(x,x')$ and not on the specific choice of clusters~$\cluster$ and~$\cluster'$ containing~$x$ and~$x'$ respectively and such that~${\cluster\setminus\{x\}=\cluster'\setminus\{x'\}}$.
\end{corollary}

@@

\begin{proof}
Let~$T\in\cat$ be a basic cluster tilting object such that~$\CC(T)$ is the initial cluster.
By~\cref{lemma:positive mutation triangles}, the morphism~$h$ belongs to the ideal~$(\susp T)$.
Therefore, the triangle ${X\to \bigoplus_i E_i \to Y \xrightarrow{h} \susp X}$ induces the relation~$[X]+[Y]=\sum_i [E_i]$ in~$K_0(\cat ; T)$.
\cref{thm::categorification} (ii) gives the equality~$\b{g}_{\CC(X)} + \b{g}_{\CC(Y)} = \sum_i\b{g}_{\CC(E_i)}$, where the~$\b{g}$-vectors are computed with respect to the initial seed~$\CC(T)$.
By \cref{thm::categorification} (iii), the middle term of the triangle is uniquely defined, up to isomorphism, by~$X$, and~$Y$, and the triangle is therefore an exchange triangle, with respect to any mutation at~$X$: it does not depend on the choice of a cluster tilting object containing~$X$.
By \cref{thm::categorification} (iv) this equality is one of the two equalities in \cref{lem:linearDependencegvectorsCA}.
\end{proof}

@@

\para{Setting for Section~\ref{sect::prelim on extricats}}
In that section, we let~$\cat$ be an extriangulated category with a fixed full additive subcategory~$\tc$, stable under isomorphisms, under taking direct summands, and satisfying the following three properties:
\begin{enumerate}
\item Every~$T\in\tc$ is projective in~$\cat$;
\item For each~$T\in\tc$, the morphism~$T\to 0$ is an inflation for the extriangulated structure of~$\cat$;
\item For each~$X\in\cat$, there is an extriangle~$T_1^X\infl T_0^X \defl X \overset{\delta_X}{\dashrightarrow}$ in~$\cat$ with~$T_0^X, T_1^X$ in~$\tc$.
\end{enumerate}

\para{Setting for Section~\ref{sect::Statment extricats}}
In that section, we keep the previous setting, but assume moreover that~$\cat$ is Krull--Schmidt, $\field$-linear, Ext-finite, and has Auslander--Reiten--Serre duality, and that the subcategory~$\tc$ is of the form~$\add T$, where~$T=T_1\oplus\cdots\oplus T_n$ is a basic object.

@@

An extriangulated category is the data of an additive category~$\cat$, an additive bifunctor~$\mathbb{E}:\cat^\mathrm{op}\times\cat\to \mathit{Ab}$ modelling the~$\operatorname{Ext}^1$-bifunctor, and an additive realization~$\mathfrak{s}$ sending each element~$\delta\in\mathbb{E}(Z,X)$ to some (equivalence classe of) diagram~$X\to Y\to Z$ modelling the short exact sequences or triangles.
Some axioms, inspired from the case of extension-closed subcategories of triangulated categories have to be satisfied.

@@

\begin{proposition}[{\cite[Prop.~2.5]{IyamaNakaokaPalu}}]
For any non-split $\mathbb{E}$-extension $\delta\in\mathbb{E}(Z,X)$, the following holds.
\begin{enumerate}[{\rm (1)}]
\item If $\delta$ satisfies {\rm (AS1)}, then $X$ is endo-local.
\item If $\delta$ satisfies {\rm (AS2)}, then $Z$ is endo-local.
\end{enumerate}
\end{proposition}

@@

\begin{lemma}
\label{lemma::first consequences}
The following holds:
\begin{enumerate}[{\rm (a)}]
\item An object belongs to~$\tc$ if and only if it is projective in~$\cat$.
\item For any~$T,T'\in\tc$, we have~$\cat(T,\susp T')=0$.
\item For any~$X\in\cat$ and~$T\in\tc$,~$\cat(X,\susp T)\cong\mathbb{E}(X,T)$.
\item For any~$X\in\cat$ and~$T\in\tc$,~$\cat(T,X)\cong\mathbb{E}(\susp T,X)$.
\item The full subcategory~$\susp\tc$ is rigid in~$\cat$.
\end{enumerate}
\end{lemma}

\begin{proof}
(a) If~$X$ is projective in~$\cat$, then the deflation~$T_0^X\defl X$ splits and~$X$ is a summand of~$T_0^X$.

(b) This follows from the fact that~$T$ is projective and~$0\defl\susp T'$ is a deflation.

(c) and (d) are obtained from the exact sequences of \cref{proposition::extricat long exact sequences} associated with the extriangle~$T\infl 0\defl \susp T\dashrightarrow$.

(e) For any~$T,T'\in\tc$, we have~$\mathbb{E}(\susp T,\susp T') \cong \cat(T,\susp T')=0$.
\end{proof}

@@

\begin{proof}[Proof of \cref{prop::KRextricat}]
Recall that~$F:\cat\to\modt$ denotes the functor sending an object~$X$ to the module~$\cat(-,X)|_\tc$.
By \cref{lemma::first consequences} (e), the ideal~$(\susp\tc)$ of morphisms factoring through an object in~$\susp\tc$ is included in the kernel of~$F$.
Conversely, let~$X\xrightarrow{f}Y$ be such that~$Ff=0$.
By \cref{lemma::shifts of extriangles}, there is a conflation~$T_0\overset{g}{\infl} X\overset{h}{\defl} \susp T_1$ with~$T_0,T_1\in\tc$.
By assumption, the composition~$f\circ g$ vanishes, which implies that~$f$ factors through~$\susp T_1$.
The functor~$F$ thus induces a faithful functor, still denoted~$F$, from~$\cat/(\susp\tc)$ to~$\modt$.
It is well-known (see for instance~\cite[Lem.~3.1]{AssemSimsonSkowronski}) that~$F$ restricts to an equivalence of categories~$\tc\to\proj\tc$.
Because~$\tc$ consists of projective objects in~$\cat$, the image under~$F$ of every conflation is a right exact sequence.
The fact that~$F$ is full and dense can thus be proven as in the case of triangulated categories, by using \cref{lemma:: mphs in T are inflations}.
\end{proof}

@@

@@

\subsubsection{Associative algebras}

>>>>

The associahedron is a ``mythical polytope''~\cite{Haiman} whose facial structure encodes Catalan families: its vertices correspond to parenthesizations of a non-associative product, triangulations of a convex polygon, or binary trees; its edges correspond to applications of the associativity rule, diagonal flips, or edge rotations; and in general its faces correspond to partial parenthesizations, diagonal dissections, or Schr\"oder trees.
Its combinatorial and topological structure was introduced in early works of D.~Tamari~\cite{Tamari} and J.~Stasheff~\cite{Stasheff}, and a first $3$-dimensional polytopal model was realized by J.~Milnor for the PhD defense of J.~Stasheff.
The first systematic polytopal realizations were constructed by M.~Haiman~\cite{Haiman} and C.~Lee~\cite{Lee}.
Since then, the associahedron has been largely ``demystified'' with several polytopal constructions and generalizations (some are discussed below).
Its realizations can be classified into three species: the secondary polytopes~\cite{GelfandKapranovZelevinsky, BilleraFillimanSturmfels}, the $\b{g}$-vectors realizations~\cite{Loday, HohlwegLange, LangePilaud, HohlwegLangeThomas, HohlwegPilaudStella, Postnikov, PilaudSantos-brickPolytope, PilaudStump-brickPolytope} and the $\b{d}$-vector realizations~\cite{ChapotonFominZelevinsky, CeballosSantosZiegler}. See~\cite{CeballosSantosZiegler} for a discussion of some of these realizations and their connections.

@@

This paper focuses on a surprising construction of the associahedron that recently appeared in mathematical physics.
Motivated by the prediction of the behavior of scattering particules, N.~Arkani-Hamed, Y.~Bai, S.~He, and G.~Yan recently described in~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} the kinetic associahedron.
It is a class of polytopal realizations of the classical associahedron obtained as sections of a high-dimensional positive orthant with well-chosen affine subspaces.
This construction provides a large degree of freedom in the choice of the parameters defining these affine subspaces, and actually produces all polytopes whose normal fan is affinely equivalent to that of J.-L.~Loday's associahedron~\cite{Loday} (see \cref{subsec:typeConeAsso}).
These realizations were then extended by V.~Bazier-Matte, G.~Douville, K.~Mousavand, H.~Thomas and E.~Y\i ld\i r\i m~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the context of finite type cluster algebras using tools from representation theory of quivers.
More precisely, they fix a finite type cluster algebra~$\cal{A}$ and consider the real euclidean space~$\R^\cal{V}$ indexed by the set~$\cal{V}$ of cluster variables of~$\cal{A}$.
Starting from an acyclic initial seed~$\seed_\circ$, they consider a set~$\cal{M}$ of mesh mutations and compute the corresponding linear dependences between the $\b{g}$-vectors, which can be interpreted as linear spaces in~$\R^\cal{V}$.
Finally, they perturb these mesh linear spaces by a collection~$\b{\ell} \in \R_{>0}^{\cal{M}}$ of positive parameters and intersect the resulting perturbed mesh affine spaces with the positive orthant~$\R_{\ge 0}^\cal{V}$ (see \cref{subsec:typeConeCA}).
They show that the resulting polytope is always a generalized associahedron, and that its normal fan is the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$.
An implicit by-product of their construction is that the space of all polytopal realizations of the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$ is a simplicial cone.

\medskip
In this paper, we revisit, extend and explore further this construction using a reversed approach.
Given a complete simplicial fan~$\Fan$, we consider the space~$\typeCone(\Fan)$ of all its polytopal realizations.
This space was called type cone in~\cite{McMullen-typeCone} and deformation cone in~\cite{Postnikov, PostnikovReinerWilliams}, who studied the case when~$\Fan$ is the braid arrangement leading to the rich theory of generalized permutahedra (recently extended to all finite Coxeter arrangements in~\cite{ACEP-DeformationsCoxeterPermutahedra}).
The type cone is known to be a polyhedral cone defined by a collection of inequalities corresponding to the linear dependences among the rays of~$\Fan$ contained in pairs of adjacent maximal cones of~$\Fan$ (see \cref{def:typeCone}).
Our approach is based on an elementary but powerful observation: for any fan~$\Fan$, all polytopal realizations of~$\Fan$ can be described as sections of a high dimensional positive orthant with a collection of affine subspaces parametrized by the type cone~$\typeCone(\Fan)$ (see \cref{prop:alternativePolytopalRealization}); if moreover the type cone~$\typeCone(\Fan)$ is a simplicial cone, it leads to a simple parametrization of all polytopal realizations of~$\Fan$ by a positive orthant corresponding to the facets of the type cone~$\typeCone(\Fan)$ (see \cref{coro:simplicialTypeCone}).
To prove that the type cone~$\typeCone(\Fan)$ is simplicial, we just need to identify which pairs of adjacent maximal cones of~$\Fan$ correspond to the facets of~$\typeCone(\Fan)$ and to show that the corresponding linear dependences among their rays positively span the linear dependence among the rays of any pair of adjacent maximal cones of~$\Fan$.
When applied to the $\b{g}$-vector fans of cluster algebras (see \cref{subsec:typeConeCA}), this reversed approach yields all polytopal realizations of the $\b{g}$-vector fans and thus efficiently revisits and extends the construction of~\cite{BazierMatteDouvilleMousavandThomasYildirim}.
This new perspective has several advantages, as our proof uniformly~applies~to:

@@

In this section, we apply the results of \cref{subsect::Statement cluster cats} in order to prove \cref{prop:uniqueExchangePropertyCA,prop:meshMutations} of \cref{subsec:typeConeCA}.

We first recall the results on categorification of cluster algebras that we will use in the proofs.
\begin{theorem}
\label{thm::categorification}
Let~$\cat$ be the cluster category of a valued quiver of Dynkin type~$A,B,C,D,E,F$ or~$G$. Let~$T$ be a basic cluster-tilting object in~$\cat$, and let~$Q$ be the valued Gabriel quiver of~$\End{\cat}(T)$. Then there is a bijection
\[
\CC: \ind(\cat) \xrightarrow{} \{\textrm{cluster variables in the cluster algebra of~$Q$}\},
\]
expressing cluster variables as Laurent polynomials in the variables of the cluster~$\CC(T)$ and which has the follwing properties.
\begin{itemize}
\item[(i)] \emph{(}\cite[Prop.~3.2]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are compatible if and only if~$\cat(X, \susp Y) = 0$.
\item[(ii)] \emph{(}\cite[Prop.~4.3]{FuKeller} and \cite[Prop.~2.2]{Palu}, where the proofs also work over an arbitrary base field\emph{)} For any~$X \in \ind(\cat)$, the~$\b{g}$-vector of~$\CC(X)$ is~$\b{g}([X])$, where we identify~$\Z^n$ and~$K_0(\cat ; T)$ via the isomorphism sending~$(a_1, \dots, a_n)$ to~$\sum_{i=1}^n a_i [T_i]$.
\item[(iii)] \emph{(}\cite[Thm.~7.5]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are exchangeable if and only if~$\dim_{\field_X}\cat(X, \susp Y) = \dim_{\field_Y}\cat(X, \susp Y) = 1$.
\item[(iv)]\emph{(}\cite{BuanIyamaReitenScott} and \cite[Thm.~5.1]{BuanMarshReiten-mutation}, adapted to the cluster categories of species of Dynkin type\emph{)} Let~$T'$ be any basic cluster-tilting object of~$\cat$. Then the matrix~$B'$ of the associated seed~$(\CC(T'),B')$ is given by the multiplicities of the middle terms of the exchange triangles. 
\end{itemize}
\end{theorem}

@@

\begin{proof}
\enlargethispage{.3cm}
If~$(x,x')$ is a positive mesh mutation, then~$\{x,x'\}$ is in~$\meshes$.  Indeed, if~$X$ and~$X'$ are the corresponding objects in~$\cat$, then there is an almost-split triangle
\[
X\to E\to X' \to \susp E,
\]
so that~$X' = \tau^{-1}X$.  Moreover, there is a slice of the Auslander--Reiten quiver of~$\cat$ containing~$X$ and all indecomposable direct factors of~$E$.  The direct sum of indecomposable objects in this slice is a cluster-tilting object in which~$X$ is a source, and mutation at~$X$ gives~$X'$.  Thus~$\{x,x'\}$ is in~$\meshes$.  This gives an injective map from the set of positive mesh mutations to~$\meshes$.

\smallskip
Next, assume that~$\{x,x'\}\in \meshes$ starts in~$x$.  Then there is a cluster-tilting object~$T$ in~$\cat$ having~$X$ as a direct factor, and such that~$X$ is a source in~$T$ and mutation at~$X$ yields~$X'$.  Since~$X$ is a source, the mutation triangle starting and ending in~$X$ are
\[
X\to E\to X' \to \susp X \quad \textrm{and} \quad  X' \to 0 \to X \to \susp X'.
\]
Therefore~$X' = \susp^{-1}X = \tau^{-1}X$.  Thus the first triangle is the almost-split triangle starting in~$X$, since the dimension of~$\cat(X',\susp X)$ is~$1$ over~$\field_{X'}$.  Since the pair~$\{x,x'\}$ is not initial,~$\phi(X')$ is not an initial variable; in other words,~$X'$ is not in~$\add(T)$.  Therefore,~$X = \tau X' = \susp X'$ is not in~$\add\susp T$, so the mutation is a positive mesh mutation by \cref{lemm::almost-split}.  This gives an injective map from~$\meshes$ to the set of positive mutations. 
\end{proof}

\begin{corollary}[\cref{lem:bijectionMeshMutations}]
\label{coro::number of positive meshes}
We have that~$|\meshes|$ is the number~$|\variables[\B_\circ]|-n$ of non-initial cluster variables.
\end{corollary}

@@

\begin{proof}
In view of \cref{thm::categorification}\,(iii), it follows from~\cite[Lem.~3.3]{Palu}.
\end{proof}

@@

\begin{proof}
Let~$\cat$ be a cluster category of Dynkin type~$A,B,C,D,E,F$ or~$G$, and let $T\in\cat$ be some basic cluster-tilting object categorifying the given initial exchange matrix~$\B_\circ$. 
By \cref{lemma:positive mutation triangles}, we may assume that the mutation under consideration is a positive mutation.
In view of \cref{thm::categorification}\,(ii) and (iv) it is enough to show that the relation in~$K_0(\cat ; T)$ given by any triangle ${X\to E\to Y\xrightarrow{h} \susp X}$ with~$h\in(\susp T)$ is a positive linear combination of relations coming from Auslander--Reiten triangles with third term not in~$\add(\susp T)$.
This categorified statement is precisely \cref{corollary::meshes positively generate cluster cats}.
\end{proof}

It also follows from \cref{thm::categorification} above that any cluster algebra of finite type has the unique exchange relation property (see \cref{def:uerp}).

\begin{corollary}
\label{corollary:UERPforCAproof}
\emph{(\cref{prop:uniqueExchangePropertyCA})}
Let~$\B_\circ$ be any finite type exchange matrix, and let~$\mathcal{A}(\B_\circ)$ be the associated cluster algebra without coefficients.
Then, for any exchangeable cluster variables~$x$ and~$x'$, the linear dependence $\sum_{y \in \cluster \cup \cluster'} \coefficient[{\b{g}_y}][\cluster][\cluster'] \, \b{g}_y = 0$ only depends on the pair~$(x,x')$ and not on the specific choice of clusters~$\cluster$ and~$\cluster'$ containing~$x$ and~$x'$ respectively and such that~${\cluster\setminus\{x\}=\cluster'\setminus\{x'\}}$.
\end{corollary}

@@

\begin{proof}
Let~$T\in\cat$ be a basic cluster tilting object such that~$\CC(T)$ is the initial cluster.
According to~\cref{lemma:positive mutation triangles}, the morphism~$h$ belongs to the ideal~$(\susp T)$.
Therefore, the triangle ${X\to \bigoplus_i E_i \to Y \xrightarrow{h} \susp X}$ induces the relation~$[X]+[Y]=\sum_i [E_i]$ in~$K_0(\cat ; T)$.
\cref{thm::categorification}\,(ii) gives the equality ${\b{g}_{\CC(X)} + \b{g}_{\CC(Y)} = \sum_i\b{g}_{\CC(E_i)}}$, where the~$\b{g}$-vectors are computed with respect to the initial seed~$\CC(T)$.
By \cref{thm::categorification}\,(iii), the middle term of the triangle is uniquely defined, up to isomorphism, by~$X$, and~$Y$, and the triangle is therefore an exchange triangle, with respect to any mutation at~$X$: it does not depend on the choice of a cluster tilting object containing~$X$.
By \cref{thm::categorification}\,(iv) this equality is one of the two equalities in \cref{lem:linearDependencegvectorsCA}.
\end{proof}

@@

\smallskip
\para{Setting for Section~\ref{sect::prelim on extricats}}
In that section, we let~$\cat$ be an extriangulated category with a fixed full additive subcategory~$\tc$, stable under isomorphisms, under taking direct summands, and satisfying the following three properties:
\begin{enumerate}
\item Every~$T\in\tc$ is projective in~$\cat$;
\item For each~$T\in\tc$, the morphism~$T\to 0$ is an inflation for the extriangulated structure of~$\cat$;
\item For each~$X\in\cat$, there is an extriangle~$T_1^X\infl T_0^X \defl X \overset{\delta_X}{\dashrightarrow}$ in~$\cat$ with~$T_0^X, T_1^X$ in~$\tc$.
\end{enumerate}

\smallskip
\para{Setting for Section~\ref{sect::Statment extricats}}
In that section, we keep the previous setting, but assume moreover that~$\cat$ is Krull--Schmidt, $\field$-linear, Ext-finite, and has Auslander--Reiten--Serre duality, and that the subcategory~$\tc$ is of the form~$\add T$, where~$T=T_1\oplus\cdots\oplus T_n$ is a basic object.

@@

An extriangulated category is the data of an additive category~$\cat$, of an additive bifunctor ${\mathbb{E}:\cat^\mathrm{op}\times\cat\to \mathit{Ab}}$ modelling the~$\operatorname{Ext}^1$-bifunctor, and of an additive realization~$\mathfrak{s}$ sending each element~$\delta\in\mathbb{E}(Z,X)$ to some (equivalence classe of) diagram~$X\to Y\to Z$ modelling the short exact sequences or triangles.
Some axioms, inspired from the case of extension-closed subcategories of triangulated categories have to be satisfied.

@@

\begin{proposition}[{\cite[Prop.~2.5]{IyamaNakaokaPalu}}]
For any non-split $\mathbb{E}$-extension $\delta\in\mathbb{E}(Z,X)$, the following holds.
\begin{enumerate}
\item If $\delta$ satisfies {\rm (AS1)}, then $X$ is endo-local.
\item If $\delta$ satisfies {\rm (AS2)}, then $Z$ is endo-local.
\end{enumerate}
\end{proposition}

@@

\begin{lemma}
\label{lemma::first consequences}
The following holds:
\begin{enumerate}[(i)]
\item An object belongs to~$\tc$ if and only if it is projective in~$\cat$.
\item For any~$T,T'\in\tc$, we have~$\cat(T,\susp T')=0$.
\item For any~$X\in\cat$ and~$T\in\tc$,~$\cat(X,\susp T)\cong\mathbb{E}(X,T)$.
\item For any~$X\in\cat$ and~$T\in\tc$,~$\cat(T,X)\cong\mathbb{E}(\susp T,X)$.
\item The full subcategory~$\susp\tc$ is rigid in~$\cat$.
\end{enumerate}
\end{lemma}

\begin{proof}
\begin{enumerate}[(i)]
\item If~$X$ is projective in~$\cat$, then the deflation~$T_0^X\defl X$ splits and~$X$ is a summand of~$T_0^X$.
\item This follows from the fact that~$T$ is projective and~$0\defl\susp T'$ is a deflation.
\item and (iv) are obtained from the exact sequences of \cref{proposition::extricat long exact sequences} associated with the extriangle~$T\infl 0\defl \susp T\dashrightarrow$.
\addtocounter{enumi}{1}
\item For any~$T,T'\in\tc$, we have~$\mathbb{E}(\susp T,\susp T') \cong \cat(T,\susp T')=0$.
\qedhere
\end{enumerate}
\end{proof}

@@

\begin{proof}[Proof of \cref{prop::KRextricat}]
Recall that~$F:\cat\to\modt$ denotes the functor sending an object~$X$ to the module~$\cat(-,X)|_\tc$.
By \cref{lemma::first consequences}\,(v), the ideal~$(\susp\tc)$ of morphisms factoring through an object in~$\susp\tc$ is included in the kernel of~$F$.
Conversely, let~$X\xrightarrow{f}Y$ be such that~$Ff=0$.
By \cref{lemma::shifts of extriangles}, there is a conflation~$T_0\overset{g}{\infl} X\overset{h}{\defl} \susp T_1$ with~$T_0,T_1\in\tc$.
By assumption, the composition~$f\circ g$ vanishes, which implies that~$f$ factors through~$\susp T_1$.
The functor~$F$ thus induces a faithful functor, still denoted~$F$, from~$\cat/(\susp\tc)$ to~$\modt$.
It is well-known (see for instance~\cite[Lem.~3.1]{AssemSimsonSkowronski}) that~$F$ restricts to an equivalence of categories~$\tc\to\proj\tc$.
Because~$\tc$ consists of projective objects in~$\cat$, the image under~$F$ of every conflation is a right exact sequence.
The fact that~$F$ is full and dense can thus be proven as in the case of triangulated categories, by using \cref{lemma:: mphs in T are inflations}.
\end{proof}

@@

We conclude with a couple of examples for two associative algebras.

@@

%\subsubsection{Associative algebras}
